A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-23 latex/main.tex
More work on ripples

See t he
\href{https://github.com/jeromekelleher/sc2ts-paper/blob/main/notebooks/fig\_mutational\_spectra.ipynb}{fig\_mutational\_spectra.ipynb}
notebook for details.

@@

\subsubsection*{Comparison with UShER+RIPPLES}
We began with the Viridian UShER tree downloaded from FigShare.
To simplify the comparison with sc2ts, we subset this 
file down to the 2,475,418 samples shared with the sc2ts ARG 
by running \texttt{matUtils extract}.
We then ran \texttt{ripples-fast}~\cite{Smith2023tracking}
on this file, using \texttt{-n1} so that all detected 
recombinants were reported (equivalent to sc2ts). We ran
ripples with the parsimony improvement parameter $p$
equal to 3 (the default) and 4.

The output of RIPPLES is two text files,

To compare these results with the sc2ts ARG, we first matched clades
by finding the most recent common ancestor (MRCA) of the samples
descending from each RIPPLES event. Overall, clade agreement was
high with [FIXME, get exact numbers] 37XX nodes (90\%) in the sc2ts ARG
where descendants match exactly. Of these nodes, 68\% had no recombinant
ancestry. Of the remainder, 595 could be matched with
one of the 855 recombination events in the sc2ts ARG

>>>>

See the
\href{https://github.com/jeromekelleher/sc2ts-paper/blob/main/notebooks/fig\_mutational\_spectra.ipynb}{fig\_mutational\_spectra.ipynb}
notebook for details.

@@

\subsubsection*{Comparison with UShER+RIPPLES}
We began with the Viridian UShER tree downloaded from FigShare.
To simplify the comparison with sc2ts, we subset this 
file down to the 2,475,418 samples shared with the sc2ts ARG 
by running \texttt{matUtils extract}.
We then ran \texttt{ripples-fast}~\cite{Smith2023tracking}
on this file, using \texttt{-n1} so that all detected 
recombinants were reported (equivalent to sc2ts). We ran
RIPPLES with the parsimony improvement parameter $p$
equal to 3 (the default) and 4.

The output of RIPPLES is two text files, ``recombination.tsv``
which lists details of putative recombination events and ``descendants.tsv``
which lists the descendant samples of all recombination events.

To compare these results with the sc2ts ARG, we first matched clades
bynding the most recent common ancestor (MRCA) of the samples
descending from each RIPPLES event. Overall, clade agreement was
high with [FIXME, get exact numbers] 37XX nodes (90\%) in the sc2ts ARG
where descendants match exactly. Of these nodes, 68\% had no recombinant
ancestry. Of the remainder, 595 could be matched with
one of the 855 recombination events in the sc2ts ARG
